An in-depth study of Explainability and Adversarial Robustness for RNNs

### 2019-12-09 ex3/paper/paper.tex
Added a reference

We think that the ``ideal'' feature importance method for our scenario are Shapley values \cite{shapley_value_1953}: To analyze feature importance with them, the feature under investigation is completely omitted from the dataset and a model is trained without it. The resulting accuracy is then compared to the baseline accuracy with all features. The effort required to do this is high both in terms of storage space as well as training time since for $n$ features $n$ models need to be separately trained and stored. While this seems tractable, when also analyzing the feature importance of combinations of features, the number of models required becomes even larger: Out of $n$ features, $2^n$ unique subsets can be chosen, meaning that the computational demand rises exponentially.

Considering the difficulties that arise from computing Shapley values, we propose a light-weight version of them, called \textit{feature dropout}: When training a model, for each sample, we leave out each feature with a certain probability by setting it to zero. On average, one feature gets zeroed out but it is also possible that none or more than one are left out. This procedure is equivalent to using dropout \cite{srivastava_dropout:_2014} with probability $\frac{1}{n}$ before the first layer. The result is the classifier learning to deal with missing features. As it can also happen that several features are left out for a sample, our method can also be used to analyze the effect of multiple features missing and can thus show possibly correlated features. The results in Table~\ref{fig:featDropout} show that, unlike the randomization feature importance method, the feature dropout method does not vastly overestimate features' importance. With feature dropout, it becomes apparent that there are very few features that actually contain unique information and that affect accuracy when left out: the destination port and the source port. A model trained with feature dropout typically yields slightly lower accuracy than a regularly trained model, even if no features are left out (flow accuracy of 0.9943 vs.~0.9965. We thus recommend training two models: One regular one and one with feature dropout to use for the feature importance.

>>>>

We think that the ``ideal'' feature importance method for our scenario are Shapley values \cite{shapley_value_1952}: To analyze feature importance with them, the feature under investigation is completely omitted from the dataset and a model is trained without it. The resulting accuracy is then compared to the baseline accuracy with all features. The effort required to do this is high both in terms of storage space as well as training time since for $n$ features $n$ models need to be separately trained and stored. While this seems tractable, when also analyzing the feature importance of combinations of features, the number of models required becomes even larger: Out of $n$ features, $2^n$ unique subsets can be chosen, meaning that the computational demand rises exponentially.

\todo{Add SHAP}

Considering the difficulties that arise from computing Shapley values, we propose a light-weight version of them, called \textit{feature dropout}: When training a model, for each sample, we leave out each feature with a certain probability by setting it to zero. On average, one feature gets zeroed out but it is also possible that none or more than one are left out. This procedure is equivalent to using dropout \cite{srivastava_dropout:_2014} with probability $\frac{1}{n}$ before the first layer. The result is the classifier learning to deal with missing features. As it can also happen that several features are left out for a sample, our method can also be used to analyze the effect of multiple features missing and can thus show possibly correlated features. The results in Table~\ref{fig:featDropout} show that, unlike the randomization feature importance method, the feature dropout method does not vastly overestimate features' importance. With feature dropout, it becomes apparent that there are very few features that actually contain unique information and that affect accuracy when left out: the destination port and the source port. A model trained with feature dropout typically yields slightly lower accuracy than a regularly trained model, even if no features are left out (flow accuracy of 0.9943 vs.~0.9965. We thus recommend training two models: One regular one and one with feature dropout to use for the feature importance. Another method that uses dropout for feature importance is \cite{chang_dropout_2017}, which, however, applies a technique called \textit{Variational Dropout} to learn the optimal dropout rate for each feature. It tries to leave out as many features as possible and at the same time keep accuracy high. Thus important features are going to be left out less often and one can then extract the dropout probabilities for each feature and assess their significance based on them.
